---
title: "Introduction to Data Science"
subtitle: "Session 10.2"
author: "Robert Clements"
date: ""
output:
  xaringan::moon_reader:
    css: rc_css.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.align = "center", fig.asp=.5, message = FALSE, warning = FALSE)
```
### Session 10.2 Outline

- Shiny apps
---
class: inverse, center, middle
# Shiny apps
---
### What is Shiny?

`shiny` is an R package that makes it easy to build interactive web apps straight from R. You can host standalone apps on a webpage or embed them in R Markdown documents or build dashboards. You can also extend your Shiny apps with CSS themes, htmlwidgets, and JavaScript actions.

You **do not need** to know `html`, `javascript` or `CSS` in order to build these apps. You only need to know `R`.

Best way to learn is by doing some simple examples here:

[https://shiny.rstudio.com/](https://shiny.rstudio.com/)
---
### Why bother?

Why use Shiny when you can use Tableau, Spotfire, Excel, etc.?

*Because Shiny has the power of R working behind it - meaning that almost anything you can do in R, you can do in a Shiny App.*

- Complex data processing

- Data exploration

- Model fitting

- Interactive visualizations

---
### Two components of a Shiny app

- The UI (user interface) object - defines what you'll see in your browser

- The server function - the code that takes inputs and creates outputs
---
### Let's run an example


```{r, eval = FALSE}
library(shiny)

runExample('01_hello')
```

See [https://shiny.rstudio.com/articles/basics.html](https://shiny.rstudio.com/articles/basics.html) for the source code of the `01_hello` app.

---
### Now let's recreat the app from scratch

Open a new R script and type in:

```{r, eval = FALSE}
library(shiny)
ui <- fluidPage(
  # App title ----
  titlePanel("Hello Shiny!")
)

server <- function(input, output) {
  # Empty server function
}
```
---
### Run the app

```{r, eval = FALSE}
shinyApp(ui, server)
```
---
### Create the skeleton for the UI

```{r, eval = FALSE}
ui <- fluidPage(
  # App title ----
  titlePanel("Hello Shiny!"),
  
  # Sidebar layout with input and output definitions ----
  sidebarLayout(
    
    # Sidebar panel for inputs ----
    sidebarPanel(
    ),
    
    # Main panel for displaying outputs ----
    mainPanel(
    )
  )
)
```
---
### Run the app

```{r, eval = FALSE}
shinyApp(ui, server)
```
---
### Add a slider input to the UI

```{r, eval = FALSE}
ui <- fluidPage(
  # App title ----
  titlePanel("Hello Shiny!"),
  
  # Sidebar layout with input and output definitions ----
  sidebarLayout(
    
    # Sidebar panel for inputs ----
    sidebarPanel(
      # Input: Slider for the number of bins ----
      sliderInput(inputId = "bins",
                  label = "Number of bins:",
                  min = 1,
                  max = 50,
                  value = 30)
    ),
    
    # Main panel for displaying outputs ----
    mainPanel(
    )
  )
)
```
---
### Add a plot output to the UI

```{r, eval = FALSE}
ui <- fluidPage(
  # App title ----
  titlePanel("Hello Shiny!"),
  
  # Sidebar layout with input and output definitions ----
  sidebarLayout(
    
    # Sidebar panel for inputs ----
    sidebarPanel(
      # Input: Slider for the number of bins ----
      sliderInput(inputId = "bins",
                  label = "Number of bins:",
                  min = 1,
                  max = 50,
                  value = 30)
    ),
    
    # Main panel for displaying outputs ----
    mainPanel(
      # Output: Histogram ----
      plotOutput(outputId = "distPlot")
    )
  )
)
```
---
### Run the app

```{r, eval = FALSE}
shinyApp(ui, server)
```
---
### Take a look at the `ui` object

```{r, eval = FALSE}
ui
```
---
### Now let's work on the server code
```{r, eval = FALSE}
server <- function(input, output) {
  
  # Load packages
  library(tidyverse)
}
```
---
### Create our plot
```{r, eval = FALSE}
server <- function(input, output) {
  
  # Load packages
  library(tidyverse)
  
  # Histogram of the Old Faithful Geyser Data ----
  # with requested number of bins
  output$distPlot <- renderPlot({

  })
}
```
---
### Run the app

```{r, eval = FALSE}
shinyApp(ui, server)
```
---
### Use the input from the slider to define the plot
```{r, eval = FALSE}
server <- function(input, output) {
  
  # Load packages
  library(tidyverse)
  
  # Histogram of the Old Faithful Geyser Data ----
  # with requested number of bins
  output$distPlot <- renderPlot({
    faithful %>% ggplot(aes(x = waiting)) +
      geom_histogram(bins = input$bins) +
      theme_bw()
  })
}
```
---
### Run the app

```{r, eval = FALSE}
shinyApp(ui, server)
```
---
class: inverse, center, middle

# End of Session 10.2