---
title: "COMPSCIX 415.2 Homework 4"
author: "Shan Valleru"
date: "2/24/2019"
output: 
  html_document:
    self_contained: true
editor_options: 
  chunk_output_type: console
---

My Github repository for my assignments can be found at this URL: [https://github.com/svalleru/compscix-415-2-assignments](https://github.com/svalleru/compscix-415-2-assignments)

```{r chunk_name, warning=FALSE, message=FALSE}
library(tidyverse)
library(dplyr)
library(nycflights13)
```

**Section 5.6.7: #2, #4 and #6 only. Extra Credit: Do #5**

2.Come up with another approach that will give you the same output as `not_cancelled %>% count(dest)` and `not_cancelled %>% count(tailnum, wt = distance)` (without using `count()`).
```{r}
flights %>% 
  filter(!is.na(dep_delay) & !is.na(arr_delay)) %>%
  group_by(dest) %>% summarize(n = n())

flights %>% 
  filter(!is.na(dep_delay) & !is.na(arr_delay)) %>%
  group_by(tailnum) %>% summarize(n = sum(distance, na.rm = TRUE))
```

4.Look at the number of cancelled flights per day. Is there a pattern? Is the proportion of cancelled flights related to the average delay?
```{r}
delayed_cancelled <-  flights %>% group_by(month, day) %>%
  summarize(avg_dep_delay = mean(dep_delay, na.rm = TRUE),
            prop_cancelled = sum(is.na(dep_time)/n()))

cor(x = delayed_cancelled$avg_dep_delay, y = delayed_cancelled$prop_cancelled)

ggplot(data = delayed_cancelled, mapping = aes(x = avg_dep_delay, y = prop_cancelled)) +
  geom_point() +
  geom_smooth(se = FALSE)

```

The correlation coefficient is positive(`0.5516646`), which means as the avergae departure delay increases, the more likely the flight is to be cancelled. This can also be see in the scatterplot.

6.What does the `sort` argument to `count()` do. When might you use it?
```{r}
?dplyr::count
```
`sort` if `TRUE` will sort output in descending order of `n`

[Extra credit]
5.Which carrier has the worst delays? Challenge: can you disentangle the effects of bad airports vs. bad carriers? Why/why not? (Hint: think about `flights %>% group_by(carrier, dest) %>% summarise(n())`)
```{r}
carrier_delay <- flights %>% 
  group_by(carrier, dest) %>% 
  summarise(n = n(), 
            arr_delay = mean(arr_delay, na.rm = TRUE))

ggplot(data = carrier_delay, mapping = aes(x = carrier, y = arr_delay)) +
  geom_boxplot()

flights %>%
  group_by(carrier) %>%
  summarise(arr_delay = mean(arr_delay, na.rm = TRUE)) %>%
  arrange(desc(arr_delay))
```

Looking at the median arrival delay from the box plot and the mean arrival delay per carrier, it seems like carries with code `F9` has the worst delays.

**Section 10.5: #1, #2, #3 and #6 only**

1.How can you tell if an object is a tibble? (Hint: try printing `mtcars`, which is a regular data frame).
```{r}
class(mtcars)
as_tibble(mtcars)
```

2.Compare and contrast the following operations on a `data.frame` and equivalent tibble. What is different? Why might the default data frame behaviours cause you frustration?
```{r}
# Data frame
df <- data.frame(abc = 1, xyz = "a")
df$x
df[, "xyz"]
df[, c("abc", "xyz")]

# Tibble
tibl <- as_tibble(df)
tibl$x
tibl[, "xyz"]
tibl[, c("abc", "xyz")]
```

Looks like data frame is returning the column even when there's a partial match in the name but tibble does not.

3.If you have the name of a variable stored in an object, e.g. `var <- "mpg"`, how can you extract the reference variable from a tibble?
```{r}
var <- "displ"
tibl <- as_tibble(mpg)
tibl$var
tibl[var]
tibl[[var]]
```

6.What option controls how many additional column names are printed at the footer of a tibble?
`tibble.max_extra_cols`: Number of extra columns printed in reduced form. Default: `100`.

**Section 12.3.3: #2, #3 and #4 only**

2.Why does this code fail?
```{r}
# table4a %>% 
#   gather(1999, 2000, key = "year", value = "cases")
#> Error in inds_combine(.vars, ind_list): Position must be between 0 and n
?table4a
table4a %>% 
  gather("1999", "2000", key = "year", value = "cases")
```

Looks like gather is trying to access 1999th and 2000th columns of the data frame. Enclosing them with double quotes seems to fix it.

3.Why does spreading this tibble fail? How could you add a new column to fix the problem?
```{r}
people <- tribble(
  ~name,             ~key,    ~value,
  #-----------------|--------|------
  "Phillip Woods",   "age",       45,
  "Phillip Woods",   "height",   186,
  "Phillip Woods",   "age",       50,
  "Jessica Cordero", "age",       37,
  "Jessica Cordero", "height",   156
)
# spread(people, key, value)
# Error: Duplicate identifiers for rows (1, 3)
# Call `rlang::last_error()` to see a backtrace
```
I think it's failing because it's unable to identify rows 1 and 3 uniquely(same name and age). Adding a column that increases the entropy should fix it.

```{r}
people <- tribble(
  ~name,             ~key,    ~value,
  #-----------------|--------|------
  "Phillip Woods",   "age",       45,
  "Phillip Woods",   "height",   186,
  "Phillip Woods",   "age",       50,
  "Jessica Cordero", "age",       37,
  "Jessica Cordero", "height",   156
)

people <- people %>%
  group_by(name, key) %>%
  mutate(entropy = row_number())

spread(people, key, value)
```
Ref: [https://www.r-bloggers.com/workaround-for-tidyrspread-with-duplicate-row-identifiers/]

4.Tidy the simple tibble below. Do you need to spread or gather it? What are the variables?
```{r}
preg <- tribble(
  ~pregnant, ~male, ~female,
  "yes",     NA,    10,
  "no",      20,    12
)
```
```{r}
preg %>%
  gather(key = 'gender', value = 'value', 2:3, na.rm = TRUE)
```

**Section 12.4.3: #1 and #2 only**

1.What do the `extra` and `fill` arguments do in `separate()`? Experiment with the various options for the following two toy datasets.
```{r}
tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"))

tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"))
?separate
# extra	
# If sep is a character vector, this controls what happens when there are too many pieces. There are three valid options:
# "warn" (the default): emit a warning and drop extra values.
# "drop": drop any extra values without a warning.
# "merge": only splits at most length(into) times
tible <- tibble(x = c("a,b,c", "d,e,f,g", "h,i,j"))
tible %>% 
  separate(x, c("one", "two", "three"), extra = 'warn')
tible %>% 
  separate(x, c("one", "two", "three"), extra = 'merge')
tible %>% 
  separate(x, c("one", "two", "three"), extra = 'drop')

# fill	
# If sep is a character vector, this controls what happens when there are not enough pieces. There are three valid options:
# "warn" (the default): emit a warning and fill from the right
# "right": fill with missing values on the right
# "left": fill with missing values on the left
tible <- tibble(x = c("a,b,c", "d,e", "f,g,i"))
tible %>% 
  separate(x, c("one", "two", "three"), fill = 'warn')
tible %>% 
  separate(x, c("one", "two", "three"), fill = 'right')
tible %>% 
  separate(x, c("one", "two", "three"), fill = 'left')
```

2.Both `unite()` and `separate()` have a `remove` argument. What does it do? Why would you set it to `FALSE`?
```{r}
?unite
?separate
# remove	
# If TRUE, remove input columns from output data frame.
# If FALSE, keeps the old column and creates a new variable.
```

