---
title: "COMPSCIX 415.2 Homework 6"
author: "Shan Valleru"
date: "3/10/2019"
output: 
  html_document:
    self_contained: true
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

My Github repository for my assignments can be found at this URL: [https://github.com/svalleru/compscix-415-2-assignments](https://github.com/svalleru/compscix-415-2-assignments)

```{r chunk_name, warning=FALSE, message=FALSE}
library(tidyverse)
library(mosaicData)
data(Whickham)
# View(Whickham)
```

# Exercise 1
1.What variables are in this data set?
```{r}
?Whickham
```
```
outcome: survival status after 20 years: a factor with levels Alive Dead
smoker: smoking status at baseline: a factor with levels No Yes
age: age (in years) at the time of the first survey
```

2.How many observations are there and what does each represent?
```{r}
?Whickham
nrow(Whickham)
# glimpse(Whickham)
```
This dataset contains a subset of the survey sample: women who were classified as current smokers or as never having smoked. The data were synthesized from the summary description tables given in the Appleton et al al paper.

3.Create a table (use the R code below as a guide) ***and a visualization*** of the relationship between smoking status and outcome, ignoring age. What do you see? Does it make sense?
```{r eval=FALSE}
library(mosaicData)
library(tidyverse)
Whickham %>% count( _________ , _________ )
```
```{r}
Whickham %>% count(smoker, outcome)

ggplot(data = Whickham, mapping = aes(x = smoker, fill = outcome), position = "fill") +
  geom_bar() +
  scale_fill_manual(values=c("#7598d1", "#ba0d1e"))
```

There are more smokers that are alive than the smokers that are dead -- which doesn't make sense!

4.Recode the `age` variable into an ordered factor with three categories: age <= 44, age > 44 & age <= 64, and age > 64. Now, recreate visualization from above, but facet on your new age factor. What do you see? Does it make sense?
```{r}
Whickham$age_category <- case_when(
  Whickham$age <= 44 ~ '<45',
  Whickham$age > 44 & Whickham$age <= 64 ~ '45-64',
  Whickham$age > 64 ~ '>64'
)
  
Whickham$age_category <- factor(Whickham$age_category, levels = c('<45', '45-64', '>64'))
ggplot(Whickham) +
  geom_bar(mapping=aes(x=smoker, fill=outcome), position="fill") +
  scale_fill_manual(values=c("#7598d1", "#ba0d1e")) +
  facet_wrap(~ age_category, ncol = 3)
```

This plot makes more sense because more smokers dead when faceted by age-- which totally different than what was shown in previous plot. I think, this exemplifies Simpson's paradox.

# Exercise 2
1.Generate a random sample of size n = 10000 from a gamma(1,2) distribution and plot a histogram or density curve. Use the code below to help you get your sample.
```{r}
library(tidyverse)
n <- 10000
# look at ?rgamma to read about this function
?rgamma
gamma_samp <- tibble(x = rgamma(n, shape = 1, scale = 2))
ggplot(data = gamma_samp, mapping = aes(x = x)) +
  geom_density()
```

2.What is the mean and standard deviation of your sample? They should both be close to 2 because for a gamma distribution:
mean = shape x scale
variance = shape x scale^2
```{r}
mean(gamma_samp$x)
sd(gamma_samp$x)
```

3.Pretend the distribution of our ***population*** of data looks like the plot above. Now take a sample of size n = 30 from a Gamma(1,2) distribution, plot the histogram or density curve, and calculate the mean and standard deviation.
```{r}
gamma_samp_30 <- tibble(x = rgamma(30, shape = 1, scale = 2))

ggplot(data = gamma_samp_30, mapping = aes(x = x)) +
  geom_density()

mean(gamma_samp_30$x)
sd(gamma_samp_30$x)
```

4.Take a sample of size n = 30, again from the Gamma(1,2) distribution, calculate the mean, and assign it to a vector named `mean_samp`. Repeat this 10000 times!!!! The code below might help.
```{r}
# create a vector with 10000 NAs
mean_samp <- rep(NA, 10000)
# start a loop
for(i in 1:10000) {
  g_samp <- rgamma(30, shape = 1, scale = 2)
  mean_samp[i] <- mean(g_samp)
}
# Convert vector to a tibble
mean_samp <- tibble(mean_samp)
```

5.Make a histogram of your collection of means from above (`mean_samp`).
```{r}
ggplot(mean_samp, mapping = aes(x = mean_samp)) +
  geom_histogram()
```

6.Calculate the mean and standard deviation of all of your sample means.
```{r}
mean(mean_samp$mean_samp)
sd(mean_samp$mean_samp)
```

7.Did anything surprise you about your answers to #6?

It looks normal.

8.According to the Central Limit Theorem, the mean of your sampling distribution should be very close to 2, and the standard deviation of your sampling distribution should be close to σn‾√=230‾‾‾√=0.365. Repeat #4-#6, but now with a sample of size n = 300 instead. Do your results match up well with the theorem?
```{r}
# create a vector with 10000 NAs
new_mean_samp <- rep(NA, 10000)
# start a loop
for(i in 1:10000) {
  new_g_samp <- rgamma(300, shape = 1, scale = 2)
  new_mean_samp[i] <- mean(new_g_samp)
}
# Convert vector to a tibble
new_mean_samp <- tibble(new_mean_samp)

mean(new_mean_samp$new_mean_samp)
sd(new_mean_samp$new_mean_samp)
```

Yes, thay match with what theorem states.
