---
title: "Introduction to Data Science"
subtitle: "Session 9.1"
author: "Robert Clements"
date: ""
output:
  xaringan::moon_reader:
    css: rc_css.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.align = "center", fig.asp=.5, message = FALSE, warning = FALSE)
```
### Housekeeping

- Final exam (homework) is cumulative (including today's material) and will be in two parts:

  + Take home portion assigned today in place of a homework assignment (Due by Friday April 5 at noon)
  
  + In class portion will be completed in class on the last day (Apr 2), and due in class by 9:30PM. 
  
  + You will not have access to course materials on Canvas a few days after the course ends. 

- Any questions?

---
### Session 9.1 Outline

- Regression with categorical variables
- Time series
  
---
class: inverse, center, middle
# Categorical Variables
---
### How do we include categorical variables in a regression model?

$$\large \hat{f}(X) = \hat{\beta}_{0} + \hat{\beta}_{1}X_{1} + \hat{\beta}_{2}X_{2} + \ldots + \hat{\beta}_{p}X_{p}$$
---
### Encoding categorical variables

There are many ways to encode categorical variables:

- Assign each category a number (Ideal = 1, Premium = 2, etc.)

- Supervised encoding (we won't go over these)

- Dummy variables/one-hot encoding (create several 0/1 variables) 
---
### Dummy variables

Suppose $\large X_{1}$ is a categorical variable with $\large k$ categories. We create $\large k-1$ dummy variables.

For example: `cut` from the diamonds data has 5 categories: `{Ideal, Premium, Good, Very Good, Fair}`

We can create 4 dummy variables as follows:

- cut_good = 1 if cut is Good, 0 otherwise
- cut_ideal = 1 if cut is Ideal, 0 otherwise
- cut_premium = 1 if cut is Premium, 0 otherwise
- cut_very_good = 1 if cut is Very Good, 0 otherwise
---
### Dummy variables

```{r echo = FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(broom)
library(knitr)
diamonds %>% select(cut) %>% head(9) %>% data.frame()
diamonds <- diamonds %>% mutate(cut_fct = factor(cut, ordered = FALSE))
diamonds_dummys <- model.matrix(~cut_fct, diamonds) %>% as_tibble()
diamonds_dummys %>% select(-`(Intercept)`) %>% head(9) %>% data.frame()
```

---
### Dummy variables

This is what `lm()` will do to an unordered factor by default.

So that this:

$$\large \hat{Price} = \hat{\beta_{0}} + \hat{\beta_{1}} \times cut$$

becomes this:

$$\large \hat{Price} = \hat{\beta_{0}} + \hat{\beta_{1}} \times Good + \hat{\beta_{2}} \times VeryGood + \hat{\beta_{3}} \times Premium + \hat{\beta_{4}} \times Ideal$$
---
### Dummy variables

```{r}
mod_1 <- lm(price ~ cut_fct, data = diamonds)
tidy(mod_1)
```
---
### What happens when `cut = Fair`?

`Fair` is our reference category, so each of the other category's coefficients are interpreted in relation to `Fair`, and each of our dummy variables equals zero.

So this

$$\large \hat{Price} = \hat{\beta_{0}} + \hat{\beta_{1}} \times Good + \hat{\beta_{2}} \times VeryGood + \hat{\beta_{3}} \times Premium + \hat{\beta_{4}} \times Ideal$$

is equivalent to this

$$\large \hat{Price} = \hat{\beta_{0}}$$
---
### What happens when `cut = Good`?

$$\large \hat{Price} = \hat{\beta_{0}} + \hat{\beta_{1}} \times Good + \hat{\beta_{2}} \times VeryGood + \hat{\beta_{3}} \times Premium + \hat{\beta_{4}} \times Ideal$$

is equivalent to this

$$\large \hat{Price} = \hat{\beta_{0}} + \hat{\beta_{1}}$$

and we interpret $\large \hat{\beta_{1}}$ as the mean difference in `Price` between `Fair` and `Good` diamonds. 
---
class: inverse, center, middle

# Time Series Analysis
---
### Time Series Decomposition

```{r, echo = FALSE, warning=FALSE, message=FALSE}
library(forecast)
library(tidyverse)
library(tseries)
daily_data <- read_csv('../../demos/session_8/day.csv')

daily_data$Date = as.Date(daily_data$dteday)
count_ts = ts(daily_data %>% select(cnt)) 

daily_data <- daily_data %>% mutate(clean_cnt = tsclean(count_ts))
daily_data <- daily_data %>% mutate(cnt_ma7 = ma(clean_cnt, order = 7),
                                    cnt_ma30 = ma(clean_cnt, order = 30))

ggplot(daily_data, aes(Date, cnt)) + geom_line() + scale_x_date('month')  + ylab("Daily Bike Checkouts") +
            xlab("")
```
---
### Time Series Decomposition

We often break a time series down by its

- trend: the overall pattern of the series (increasing, decreasing, non-seasonal cycles)
- seasonality: fluctuations over the calendar cycle
- error/residual: everything else

$$\large Y_t = S_t + T_t + E_t$$
---
### Time Series Decomposition

We can use the `stl()` function in base R to do this.

If we think a `week` is a natural unit of time:

```{r, fig.asp = .6}
# Convert to time series object
# frequency is number of observations/time unit
cln_cnt <- ts(as.vector(daily_data$clean_cnt), frequency=7)
# decompose
decomp <- stl(cln_cnt, s.window = "periodic")
plot(decomp)
```
---
### Time Series Decomposition

We can use the `stl()` function in base R to do this.

If we think a `month` is a natural unit of time:

```{r, fig.asp = .6}
# Convert to time series object
# frequency is number of observations/time unit
cln_cnt <- ts(as.vector(daily_data$clean_cnt), frequency=30)
# decompose
decomp <- stl(cln_cnt, s.window = "periodic")
plot(decomp)
```
---
### Time Series Decomposition

We can use the `stl()` function in base R to do this.

If we think a `quarter` is a natural unit of time:

```{r, fig.asp = .6}
# Convert to time series object
# frequency is number of observations/time unit
cln_cnt <- ts(as.vector(daily_data$clean_cnt), frequency=90)
# decompose
decomp <- stl(cln_cnt, s.window = "periodic")
plot(decomp)
```
---
### Time Series Decomposition

We can use the `stl()` function in base R to do this.

If we think a `half-year` is a natural unit of time:

```{r, fig.asp = .6}
# Convert to time series object
# frequency is number of observations/time unit
cln_cnt <- ts(as.vector(daily_data$clean_cnt), frequency=180)
# decompose
decomp <- stl(cln_cnt, s.window = "periodic")
plot(decomp)
```
---
### Time Series Decomposition

We can use the `stl()` function in base R to do this.

If we think a `year` is a natural unit of time:

```{r, fig.asp = .6}
# Convert to time series object
# frequency is number of observations/time unit
cln_cnt <- ts(as.vector(daily_data$clean_cnt), frequency=360)
# decompose
decomp <- stl(cln_cnt, s.window = "periodic")
plot(decomp)
```
---
### Time Series Decomposition

```{r}
class(decomp)
names(decomp)
head(decomp$time.series)
```
---
### Time Series Decomposition

```{r echo=FALSE}
decomp_series <- as_tibble(decomp$time.series)
decomp_series <- decomp_series %>% mutate(decomp_sum = seasonal + trend + remainder, clean_cnt = as.vector(daily_data$clean_cnt))
head(decomp_series)
```
---
### Dynamic Time Series Plots

We can use the `dygraphs` R package to create interactive time series plots. The package provides functions that are wrappers around the dygraphs javascript library.

```{r}
library(dygraphs)
dygraph(data = daily_data$clean_cnt,
        xlab = 'Day',
        ylab = 'Bicycle Count')
```
---
class: inverse, center, middle

# End of Session 9.1