---
title: "Introduction to Data Science"
subtitle: "Session 5.3"
author: "Robert Clements"
date: ""
output:
  xaringan::moon_reader:
    css: rc_css.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.align = "center", fig.asp=.5, message = FALSE, warning = FALSE)
```

### Session 5.3 Outline

- EDA
  + Covariation and Correlation
  + Patterns and Models
  + *R for Data Science - Chapter 7*
---
class: inverse, center, middle
# Exploratory Data Analysis
---
### Covariation and Correlation

Covariation is the tendency for the values of two or more variables to vary together in a related way.

There are three scenarios to deal with:

1. **One categorical and one continuous variable**

2. Two categorical variables

3. Two continuous variables.

---
### Visualizing covariation

Scenario 1: use boxplots.

What is a surprising finding from this plot?

```{r, warning=FALSE, message=FALSE}
library(tidyverse)

ggplot(data = diamonds, mapping = aes(x = cut, y = price)) + 
  geom_boxplot()

```
---
### Visualizing covariation

Scenario 1: use boxplots.

Sometimes R doesn't know in advance that our categorical variable is ordinal. Or, sometimes you simply want to reorder a variable for visual reasons.

```{r}
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) +
  geom_boxplot()
```
---
### Visualizing covariation

Scenario 1: use boxplots.

Sometimes R doesn't know in advance that our categorical variable is ordinal. Or, sometimes you simply want to reorder a variable for visual reasons. Let's reorder `class` by the median of `hwy`.

```{r}
ggplot(data = mpg) +
  geom_boxplot(mapping = aes(x = reorder(class, hwy, FUN = median), y = hwy))
```
---
### Pop Quiz

Sometimes category names are large and so it would make sense to plot boxplots on the y-axis. Reproduce this plot from the `diamonds` data set:

```{r, echo = FALSE}
ggplot(data = diamonds, mapping = aes(x = reorder(cut, price, FUN = median), y = price)) + 
  geom_boxplot() +
  xlab('CUT') +
  ylab('PRICE') +
  coord_flip()
```
---
### Covariation and Correlation

Covariation is the tendency for the values of two or more variables to vary together in a related way.

There are three scenarios to deal with:

1. One categorical and one continuous variable

2. **Two categorical variables**

3. Two continuous variables.

---
### Visualizing covariation

Scenario 2: use a heatmap (not a mosaic) with added text using `geom_tile()` and `geom_text()`.

What do we notice about the color and cut of diamonds?

```{r}
diamonds %>% 
  count(color, cut) %>% 
  ggplot(mapping = aes(x = color, y = cut)) +
  geom_tile(mapping = aes(fill = n)) +
  geom_text(aes(label = n), color = 'white')
```
---
### Covariation and Correlation

Covariation is the tendency for the values of two or more variables to vary together in a related way.

There are three scenarios to deal with:

1. One categorical and one continuous variable

2. Two categorical variables

3. **Two continuous variables**.

---
### Visualizing covariation

Scenario 3: use scatterplots.

What is wrong with this plot? What are some suggestions for fixing it?

```{r}
ggplot(data = diamonds) +
  geom_point(mapping = aes(x = carat, y = price))
```
---
### Visualizing covariation with larger data sets

Scenario 3: use 2-d binning with the `geom_bin2d()` geom.

```{r}
ggplot(data = diamonds) +
  geom_bin2d(mapping = aes(x = carat, y = price))
```
---
### Patterns and models

Patterns reveal possible relationships. Ask yourself these questions when you spot a pattern:

1. Is the pattern due to random chance?

2. What does the relationship look like? Is it positive, negative, up down? Describe it.

3. How strong is the relationship? Weak? Strong? So-so?

4. Are there other variables that might affect this relationship?

5. Does the relationship change if you look at subgroups of the data?
---
### The importance of patterns

Patterns reveal covariation (or correlation). Although *variance* is a measure of uncertainty and spread, *covariance* can actually help **reduce** uncertainty. If two (or more) variables have high covariance, you can use the values of one (or more) to predict, or explain, the values of another.

In rare cases of causal relationships, you can use the value of a variable to **control** the value of another.

We can use *models* to help us extract patterns out of data.
---
### Patterns and models

Look at the pattern in this histogram of eruption times of Old Faithful in Yellowstone National Park.

Why does this pattern exist? Is it random? Is it a strong pattern?

```{r}
ggplot(data = faithful, aes(x = eruptions)) +
  geom_histogram() +
  xlab('Eruption time in minutes')
```
---
### Pop Quiz

This is a histogram of the eruption times (in minutes) of the Old Faithful geyser in Yellowstone National Park. What might explain this shape?

```{r, echo=FALSE}
ggplot(data = faithful, aes(x = eruptions)) +
  geom_histogram() +
  xlab('Eruption time in minutes')
```
---
### Pop Quiz

Look at the `faithful` data set and ask yourself: is there a variable that might affect the pattern seen below? If so, make a plot that reveals the reason for the pattern seen in this histogram. What is the reason?

```{r, echo=FALSE}
ggplot(data = faithful, aes(x = eruptions)) +
  geom_histogram() +
  xlab('Eruption time in minutes')
```
---
### Your first taste of a model

Remember this counterintuitive graphic?

How can we explain this result?

```{r, warning=FALSE, message=FALSE}
ggplot(data = diamonds, mapping = aes(x = cut, y = price)) + 
  geom_boxplot()

```
---
### Your first taste of a model

The pattern between carat and price is strong, positive, and explainable. Let's go back to our list of questions.

```{r}
ggplot(data = diamonds) +
  geom_bin2d(mapping = aes(x = carat, y = price))
```
---
### Your first taste of a model

1. Is the pattern due to random chance?
2. What does the relationship look like? Is it positive, negative, up, down? Describe it.
3. How strong is the relationship? Weak? Strong? So-so?
4. Are there other variables that might affect this relationship? **ding! ding! ding!**
5. Does the relationship change if you look at subgroups of the data?

```{r}
ggplot(data = diamonds) +
  geom_bin2d(mapping = aes(x = carat, y = price))
```
---
### Thought experiment

How might we reveal a more intuitive pattern between the price and cut of diamonds by considering the very complex relationships between price, carat, and cut? 


---
class: inverse, center, middle

# End of Session 5.3