---
title: "COMPSCIX 415.2 Homework 3"
author: "Shan Valleru"
date: "2/17/2019"
output: 
  html_document:
    self_contained: true
editor_options: 
  chunk_output_type: console
---

My Github repository for my assignments can be found at this URL: [https://github.com/svalleru/compscix-415-2-assignments](https://github.com/svalleru/compscix-415-2-assignments)

```{r chunk_name, warning=FALSE, message=FALSE}
library(tidyverse)
library(nycflights13)
```

Section 3.8.1: all exercises
```{r}
# 1. What is the problem with this plot? How could you improve it?
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point()
```

There's lot of overlap of the dots. Setting some jitter can fix that.
```{r}
mpg_gg <- ggplot(mpg, mapping = aes(x = cty, y = hwy))
mpg_gg +
  geom_jitter()
```

2. What parameters to `geom_jitter()` control the amount of jittering?
```{r}
?geom_jitter
```

`width`	and `height`

3. Compare and contrast geom_jitter() with geom_count().
```{r}
mpg_gg +
  geom_jitter()

mpg_gg +
  geom_count()
```

`geom_jitter` adds jitter to overlapping points where as `geom_count` makes the overlapping points look bigger on the plot

4. What’s the default position adjustment for `geom_boxplot()`? Create a visualisation of the `mpg` dataset that demonstrates it.
```{r}
?geom_boxplot
# Boxplots are automatically dodged when any aesthetic is a factor
# p + geom_boxplot(aes(colour = drv))

ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot(position = "dodge") +
  coord_flip()
```

Default postion adjustment is `dodge`

Section 3.9.1: #2 and #4 only

2. What does `labs()` do? Read the documentation.
```{r}
?labs
# example(labs)
```

It lets you add title, subtitle, caption, tag, label to the graphs

4. What does the plot below tell you about the relationship between city and highway mpg? Why is `coord_fixed()` important? What does `geom_abline()` do?
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline() +
  coord_fixed()
```

The plot shows that the highway mileage is better than the city mileage-- meaning, better fuel efficiency on highways that in the cities.

Section 4.4: #1 and #2 only

1. Why does this code not work?
```{r}
my_variable <- 10
# my_varıable
#> Error in eval(expr, envir, enclos): object 'my_varıable' not found
```

There's a typo in the second variable. It should've been `my_variable`

2. Tweak each of the following R commands so that they run correctly:
```{r}
library(tidyverse)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
# i don't see any issues with this!!

# fliter(mpg, cyl = 8)
# typo in the command; the following will work
filter(mpg, cyl == 8)

# filter(diamond, carat > 3)
# typo in the data name
filter(diamonds, carat > 3)
```

Section 5.2.4: #1, #3 and #4 only. You will need to install the `nycflights13` package and use the `flights` data.

1. Find all flights that
```{r}
# Had an arrival delay of two or more hours
flights %>% filter(arr_delay >= 120)

# Flew to Houston (IAH or HOU)
flights %>% filter(dest %in% c("IAH", "HOU"))

# Were operated by United, American, or Delta
flights %>% filter(carrier %in% c("UA", "AA", "DL"))

# Departed in summer (July, August, and September)
flights %>% filter(month %in% 7:9)

# Arrived more than two hours late, but didn’t leave late
flights %>% filter(arr_delay > 120 & dep_delay <= 0)

# Were delayed by at least an hour, but made up over 30 minutes in flight
flights %>% filter(dep_delay > 60 & dep_delay - arr_delay > 30)

# Departed between midnight and 6am (inclusive)
flights %>% filter(dep_time %in% 0:600)
```

3. How many flights have a missing `dep_time`? What other variables are missing? What might these rows represent?
```{r}
sum(is.na(flights$dep_time))
summary(flights)
```

variable_name missing_count

`dep_time`      `8255`

`dep_delay`     `8255`

`arr_time`      `8713`

`arr_delay`     `9430`

`air_time`      `9430`

4. Why is `NA ^ 0` not missing? Why is `NA | TRUE` not missing? Why is `FALSE & NA` not missing? Can you figure out the general rule? (`NA * 0` is a tricky counterexample!)

Anything power 0 is 1, so, `NA ^ 0` always evaluates to 1
`NA | TRUE` always evaluates to `TRUE` because at least one of the operator is `TRUE`
`FALSE & NA` always evaluates to `FALSE` because at least one of the operator is `FALSE`

Section 5.4.1: #1 and #3 only

1. Brainstorm as many ways as possible to select `dep_time`, `dep_delay`, `arr_time`, and `arr_delay` from `flights`.
```{r}
my_variables <- c("dep_time", "dep_delay", "arr_time", "arr_delay")
for (my_var in my_variables){
  print(flights %>% select(my_var))
}

for (my_var in my_variables){
  print(select(flights, my_var))
}

select(flights, one_of(my_variables))
select(flights, my_variables)

flights %>% select(my_variables)
```

3. What does the `one_of()` function do? Why might it be helpful in conjunction with this vector?
```{r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
select(flights, one_of(vars))
```

It's making `select` pick the rows that are in the vector `vars`