---
title: "Introduction to Data Science"
subtitle: "Session 8.2"
author: "Robert Clements"
date: ""
output:
  xaringan::moon_reader:
    css: rc_css.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.align = "center", fig.asp=.5, message = FALSE, warning = FALSE)
```

### Session 8.2 Outline

- Spatial Data
  + *Modern Data Science with R - Chapter 14*
---
class: inverse, center, middle
# Spatial Data
---
### Why spatial data?

Example: Cholera in 1854 - back then it was believed that cholera was airborne. A physician named John Snow didn't buy it, and decided to dig into the data. The data might have looked like this:

```{r, echo = FALSE}
chol <- data.frame(Date = c('Aug 31, 1854', 'Aug 31, 1854', 'Oct 1, 1854'),
                   Name = c('Thomas Jones', 'Mary Jones', 'Martin Warwick'),
                   Address = c('26 Broad St', '26 Broad St', '14 Broad St'),
                   Age = c(37, 11, 23), 
                   Death = c('cholera', 'cholera', 'cholera'))
chol
```
---
### Why spatial data?


```{r, echo = FALSE, warning=FALSE, message=FALSE, fig.asp = .8}
library(mdsr)
library(ggmap)
library(rgdal)
m <- get_map('John Snow, London, England', zoom = 17, maptype = 'roadmap', source = 'google', urlonly = FALSE)
proj4string(CholeraDeaths) <- CRS("+init=epsg:27700")
cholera_latlong <- CholeraDeaths %>% spTransform(CRS("+init=epsg:4326"))
snow <- ggmap(m) + geom_point(data = as.data.frame(cholera_latlong), aes(x = coords.x1, y = coords.x2, size = Count))
snow
```
---
### Spatial objects in R

There are many types of spatial objects. `CholeraDeaths` is a `SpatialPointsDataFrame` object available in the `mdsr` R package. 

```{r}
library(mdsr)
glimpse(CholeraDeaths)
```
---
### Spatial objects in R

Luckily, we can convert `CholeraDeaths` to a standard R data frame which will make plotting easier.

```{r}
cholera_df <- as.data.frame(CholeraDeaths)
head(cholera_df)
```
---
### Wait, what happened to our coordinates?

When working with spatial data, we have to worry about **projections** - how do we convert locations in a 3-d geographic coordinate system to a 2-d representation. 
.pull-left[
Mercator preserves angles
```{r, echo = FALSE, warning = FALSE, message = FALSE, fig.asp = 1}
library(maps)
map("world", projection = 'mercator', wrap = TRUE)
```
]
.pull-right[
Gall-Peters preserves area
```{r, echo = FALSE, warning = FALSE, message = FALSE, fig.asp = 1}
map("world", projection = 'cylequalarea', param = 45, wrap = TRUE)
```
]
---
### Google map tiles

Google map tiles are returned to R using the `espg:4326` system, which is the standard for GPS systems. We need to project our `CholeraDeaths` coordinates onto the same system. 

We can use the `spTransform()` and `CRS()` functions from the `rgdal` package to do this transformation. `spTransform` does the conversion, and `CRS` specifies the coordinate system using a smaller string.  

But first, our cholera data is in the `epsg:27700` system. We need to tell R this by using the `proj4string()` function. Without this, we can't convert correctly.

```{r}
library(rgdal)
# specify the system for our spatial data
proj4string(CholeraDeaths) <- CRS("+init=epsg:27700")
```
---
### Conversion
```{r}
# transform to new system
cholera_latlong <- CholeraDeaths %>% spTransform(CRS("+init=epsg:4326"))
head(cholera_latlong)
```

---
### Create a data.frame with lat and long columns
```{r}
cholera_latlong <- as.data.frame(cholera_latlong)
head(cholera_latlong)
```
---
### Create a data.frame with lat and long columns
```{r}
cholera_latlong <- cholera_latlong %>% 
  mutate(longitude = coords.x1, latitude = coords.x2)
head(cholera_latlong)
```
---
### Now we can plot it on a map

We'll use the `ggmap` package to plot static maps.

Unfortunately, we need an API key to do this. After installing `ggmap`, read the documentation by typing

```{r, eval = FALSE}
?register_google
```

**Not required**, but you can get your API key at:

[https://cloud.google.com/maps-platform/](https://cloud.google.com/maps-platform/)

---
### Now we can plot it on a map

First, we collect the map tiles using `get_map()`.
Second, we use `ggplot` to plot the map and add our spatial coordinates as points.

```{r, echo = TRUE, warning=FALSE, message=FALSE, fig.asp = .8, eval = FALSE}
library(mdsr)
library(ggmap)
library(rgdal)

# get map tiles
m <- get_map('John Snow, London, England', zoom = 17, 
             maptype = 'roadmap', source = 'google', urlonly = FALSE)

snow <- ggmap(m) + 
  geom_point(data = cholera_latlong, 
             aes(x = longitude, y = latitude, size = Count))
```
---
### Now we can plot it on a map

```{r}
snow
```
---
### Geocoding

Geocoding is the process of converting a human-readable address into geographic coordinates.

We can get latitudes and longitudes from an address by using the `geocode()` function from the `ggmap` package. 

```{r}
library(ggmap)
ucb_ext <- '160 Spear St, San Francisco, CA'
geocode(ucb_ext)
```
---
### Slippy Maps 

A slippy map is a map with a tiled base layer, with panning and zooming and other features that you supply.

```{r, echo = FALSE}
library(leaflet)
ucb_geo <- geocode(ucb_ext)
map <- leaflet() %>%
  addTiles() %>%
  addMarkers(lng = ~lon, lat = ~lat, data = ucb_geo)
map
```
---
### Slippy Maps in R

The `leaflet` R package is a wrapper for the `leaflet` javascript library. What that means is that we can use R to create slippy maps. Note that these are meant for **web maps**. You cannot interact with a map in *pdf* or *Word*. So, your homework this week will have to be turned in as an html file.

```{r}
library(leaflet)
leaflet() %>% addTiles()
```
---
### Slippy Maps in R

Let's focus on SF

```{r}
leaflet() %>% 
  setView(lng = -122.3937, lat = 37.79137, zoom = 14) %>%
  addTiles()
```
---
### Slippy Maps in R

Let's add a marker for UC Berkeley Extension

```{r}
# geocode uc berkeley extension
ucb_geo <- geocode(ucb_ext)

map <- leaflet() %>%
  addTiles() %>%
  addMarkers(lng = ~lon, lat = ~lat, data = ucb_geo)

map
```
---
### Slippy Maps for spatial data

If we have a set of latitudes and longitudes, `leaflet()` also takes a `data` argument.

```{r}
head(cholera_latlong)

```
---
### Slippy Maps for spatial data

If we have a set of latitudes and longitudes, `leaflet()` also takes a `data` argument.

```{r}
cholera <- leaflet(cholera_latlong) %>%
  addTiles() %>%
  addMarkers()
```
---
### Slippy Maps for spatial data

```{r}
cholera
```
---
### Play around with `leaflet()`

You can add labels to the markers, you can add popups (`addPopups()`). Check out the examples from the help file.

---
class: inverse, center, middle

# End of Session 8.2