---
title: "Introduction to Data Science"
subtitle: "Session 8.3"
author: "Robert Clements"
date: ""
output:
  xaringan::moon_reader:
    css: rc_css.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.align = "center", fig.asp=.5, message = FALSE, warning = FALSE)
```

### Session 8.3 Outline

- Time Series Data
---
class: inverse, center, middle
# Time Series Data
---
### Why time series data?

Many business use cases:

- Explaining seasonal patterns in sales

- Predicting the expected number of incoming or churning customers

- Estimating the effect of a newly launched product on number of sold units

- Detecting unusual events and estimating the magnitude of their effect
---
### Why time series data?

```{r, echo = FALSE, warning=FALSE, message=FALSE}
library(forecast)
library(tidyverse)
library(tseries)
daily_data <- read_csv('../../demos/session_8/day.csv')

daily_data$Date = as.Date(daily_data$dteday)

ggplot(daily_data, aes(Date, cnt)) + geom_line() + scale_x_date('month')  + ylab("Daily Bike Checkouts") +
            xlab("")
```
---
### Bike data

First, let's install a package: `forecast` for time series modeling.
Next, download the bike data (`day.csv`) and load it into `R`.

```{r, eval = FALSE}
library(forecast)
library(tidyverse)
library(tseries)

daily_data <- read_csv('path/to/day.csv')
```
---
### Bike data

```{r}
glimpse(daily_data)
```
---
### EDA

Let's convert the `dteday` column to a Date object and plot the time series for the number of bikes checked out per day.

```{r, echo = TRUE, warning=FALSE, message=FALSE}
daily_data$Date = as.Date(daily_data$dteday)

ggplot(daily_data, aes(Date, cnt)) + 
  geom_line() + 
  ylab("Daily Bike Checkouts") +
  xlab("") +
  theme_bw()
```
---
### EDA

What do we see as far as:

- trend
- seasonality
- volatility?

Do these make sense?

```{r, echo = FALSE, warning=FALSE, message=FALSE}
ggplot(daily_data, aes(Date, cnt)) + 
  geom_line() + 
  ylab("Daily Bike Checkouts") +
  xlab("") +
  theme_bw()
```
---
### Let's clean it up a little

We can remove outliers with the `tsclean()` function.

```{r}
count_ts = ts(daily_data %>% select(cnt)) 

daily_data <- daily_data %>% mutate(clean_cnt = tsclean(count_ts))

ggplot() +
  geom_line(data = daily_data, aes(x = Date, y = clean_cnt)) + 
  ylab('Cleaned Bicycle Count') +
  theme_bw()
```
---
### The Moving Average

A simple way of smoothing out a time series is to look at the moving average, i.e. the average value inside a moving window of some length of time.

$$\Huge MA = \frac{1}{m}\sum_{i=-k}^{k} y_{t+i}$$

where $\huge k$ is half of the size of the time window.
---
### The Moving Average

We can use the `ma()` function from the `forecast` package.

```{r}
daily_data <- daily_data %>% mutate(cnt_ma7 = ma(clean_cnt, order = 7),
                                    cnt_ma30 = ma(clean_cnt, order = 30))

daily_data %>% select(cnt, cnt_ma7, cnt_ma30)
```
---
### The Moving Average

```{r}
ggplot(data = daily_data) +
  geom_line(aes(x = Date, y = clean_cnt, colour = "Counts")) +
  geom_line(aes(x = Date, y = cnt_ma7,   colour = "Weekly Moving Average"))  +
  geom_line(aes(x = Date, y = cnt_ma30, colour = "Monthly Moving Average"))  +
  ylab('Bicycle Count') +
  theme_bw()
```
---
### Time Series Decomposition

We often break a time series down by its

- trend: the overall pattern of the series (increasing, decreasing, non-seasonal cycles)
- seasonality: fluctuations over the calendar cycle
- error/residual: everything else

$$\Huge Y_t = S_t + T_t + E_t$$
---
### Time Series Decomposition

We can use the `stl()` function in base R to do this.

```{r, fig.asp = .6}
# Convert to time series object
# frequency is number of observations/time unit
count_ma <- ts(na.omit(daily_data$cnt_ma30), frequency=30)

# decompose
decomp <- stl(count_ma, s.window="periodic")

plot(decomp)
```
---
### Question for you

Why do we even care about decomposing the time series? Why don't we just use linear regression to forecast new values using some features, and ignore time?
---
### Dynamic Time Series Plots

We can use the `dygraphs` R package to create interactive time series plots. The package provides functions that are wrappers around the dygraphs javascript library.

```{r}
library(dygraphs)
dygraph(data = daily_data$clean_cnt,
        xlab = 'Day',
        ylab = 'Bicycle Count')
```
---
class: inverse, center, middle

# End of Session 8.3