---
title: "Introduction to Data Science"
subtitle: "Session 8.1"
author: "Robert Clements"
date: ""
output:
  xaringan::moon_reader:
    css: rc_css.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.align = "center", fig.asp=.5, message = FALSE, warning = FALSE)
```
### Housekeeping

- Final exams (homework) will be cumulative, open note/book, and will be in two parts:

  + Take home portion will be assigned next week in place of a homework assignment (Due by Friday April 5th at noon)
  
  + In class portion will be completed in class on the last day (Apr 2), and due in class by 9:30PM. 

- Any questions?

---
### Session 8.1 Outline

- R data structures
- R functions and loops

---
class: inverse, center, middle
# R Data Structures
---
### We've been using these structures a lot:

- vectors:  all of the same type (character, numeric, etc.)
```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
```

```{r}
x <- c(1,2,3, 11, 12, 13)
x
```

- data frames: two-dimensional; columns can be of different types 

```{r}
my_df <- data.frame(x)
my_df
```
---
### We've been using these structures a lot:

- tibbles: similar to data frames

```{r}
my_tbl <- as_tibble(x)
my_tbl
```
---
### There are two other very common types: matrix and list

- matrix: two-dimensional; all same type

```{r}
mdat <- matrix(x, nrow = 2, ncol = 3, byrow = TRUE)
mdat
```

---
### There are two other very common types: matrix and list

- matrix: two-dimensional; all same type

```{r}
mdat <- matrix(x, nrow = 2, ncol = 3, byrow = FALSE)
mdat
```
---
### There are two other very common types: matrix and list

A `list` is a generic vector containing other objects.

```{r}
my_list <- list(a = c(1,2,3), 
                b = c('a', 'b', 'c'), 
                c = tibble(x = c(3,2,1), y = c('b', 'd', 'f')))
my_list
```
---
### R `list` structures

They are similar to python `dictionaries` (key-value pairs). If you want to see the keys, you can use the `names()` function.

```{r}
names(my_list)
```
---
### R `list` structures

Extract elements of a list:

- by name: `$` or `[[]]`
- by position: `[[]]`

```{r}
my_list[['a']]
my_list$a
my_list[[1]]
```
---
### Many, many functions in R return `list`-type objects

```{r}
my_lm <- lm(price ~ carat, data = diamonds)
names(my_lm)
```

---
### Many, many functions in R return `list`-type objects

```{r}
my_lm[['coefficients']]
my_lm$coefficients
my_lm[[1]]
```
---
### Tibbles with `list` columns

Remember this from last time?

```{r}
n <- 1000 #how many samples
slopes <- rep(NA, n) # empty vector for saving coefficients
for(i in 1:n) {
  diam_samp <- diamonds %>% sample_n(10000) # random sample
  diam_samp_lm <- lm(price ~ carat, data = diam_samp)
  slopes[i] <- coef(diam_samp_lm)[2] # store the coefficient
}
slopes <- as_tibble(slopes)
slopes
```
---
### Tibbles with `list` columns

Let's do the same thing, but let's save our 1000 samples, models, and slopes all in one tibble. 

```{r}
n <- 1000 #how many samples
slopes <- rep(NA, n) # empty vector for saving coefficients
my_mods <- my_samps <- list() # empty lists for saving models and samples
for(i in 1:n) {
  diam_samp <- diamonds %>% sample_n(1000) # random sample
  diam_samp_lm <- lm(price ~ carat, data = diam_samp)
  slopes[i] <- coef(diam_samp_lm)[2] # store the coefficient
  my_mods[[i]] <- diam_samp_lm
  my_samps[[i]] <- diam_samp
}
all_results <- tibble(my_samps, my_mods, slopes)
```
---
### Tibbles with `list` columns

Let's do the same thing, but let's save our 1000 samples, models, and slopes all in one tibble. 

```{r}
all_results
```
---
class: inverse, center, middle
# R Functions and Loops
---
### Defining R functions

I want to run this code over and over again, but maybe I want to be able to modify the sample size, or the x variable in my regression model - this seems like the perfect time to talk about **functions**!!

```{r, eval = FALSE}
diam_samp <- diamonds %>% sample_n(1000) # random sample
diam_samp_lm <- lm(price ~ carat, data = diam_samp)
slopes <- coef(diam_samp_lm)[2] # store the coefficient
```
---
### Defining R functions

Each function could (but not must) have inputs and outputs:

```{r, eval = FALSE}
# name your function
name_of_function <- function(input arguments) { # define input arguments 
  R code here
  ...
  return(what you want to return) # explicitly return something
}

# call the function
name_of_function(input arguments)

# or assign the output
output_name <- name_of_function(input arguments)
```
---
### Defining R functions

```{r}
fit_model <- function(samp_size) {
  diam_samp <- diamonds %>% sample_n(samp_size) # random sample
  diam_samp_lm <- lm(price ~ carat, data = diam_samp)
  slopes <- coef(diam_samp_lm)[2] # store the coefficient
  return(slopes)
}

# call function
fit_model(100)
```
---
### Defining R functions

Let's do the same for our regression model formula. We want to be able to pass to the function the name of the predictor variable: `price ~ ____`. 

Easiest way to help us do this is to use the `paste0` function.

```{r}
predictor <- 'carat'
mod_formula <- paste0('price ~ ', predictor)
mod_formula
```
---
### Pop Quiz

Write a function that takes two arguments: (1) samp_size - sample size; (2) x_var - predictor variable. This function should take a sample from the diamonds data set of size `samp_size`, fit a regression model predicting `price` from `x_var`, and then return the slope coefficient from that model.

```{r, echo = FALSE, eval=FALSE}
fit_model <- function(samp_size, x_var) {
  diam_samp <- diamonds %>% sample_n(samp_size) # random sample
  diam_samp_lm <- lm(paste0('price ~ ', x_var), data = diam_samp)
  slopes <- coef(diam_samp_lm)[2] # store the coefficient
  return(slopes)
}

# call function
fit_model(100, 'carat')
```
---
### Pop Quiz

Modify your function so that it returns not just the slopes, but also the model and the sample of data, in a list.

```{r, echo = FALSE, eval=FALSE}
fit_model <- function(samp_size, x_var) {
  diam_samp <- diamonds %>% sample_n(samp_size) # random sample
  diam_samp_lm <- lm(paste0('price ~ ', x_var), data = diam_samp)
  slopes <- coef(diam_samp_lm)[2] # store the coefficient
  all_results <- list(my_samp = diam_samp, my_mod = diam_samp_lm, slope = slopes)
  return(all_results)
}
```
---
### R `for` loops

I want to call this function over and over again and save all of the results - this seems like the perfect time to talk about **loops**!

We will only talk about `for` loops
```{r, eval = FALSE}
# create space for the output of your loop 
output_vector <- rep(NA, length of loop)

# start the loop
for(index in some sequence of values) {
  R code here
  ...
  output_vector[index] <- output # save the output for each iteration
}
```
---
### Pop Quiz

Using a `for` loop, run the function we just created 3 times, use a sample of size 100 and `carat` as our predictor, save the results, and then put them in a tibble.

```{r, eval = FALSE}
n <- 3
slopes <- rep(NA, n) # empty vector for saving coefficients
my_mods <- my_samps <- list() # empty lists for saving models and samples
for(i in 1:n) {
  my_result <- fit_model(100, 'carat')
  
  # extract the slope and store it in slopes
  # store the model in my_mods
  # store the sample in my_samps
}
all_results <- tibble(my_samps, my_mods, slopes)
```

```{r, echo = FALSE, eval = FALSE}
n <- 3
slopes <- rep(NA, n) # empty vector for saving coefficients
my_mods <- my_samps <- list() # empty lists for saving models and samples
for(i in 1:n) {
  my_result <- fit_model(100, 'carat')
  slopes[i] <- my_result$slope # store the coefficient
  my_mods[[i]] <- my_result$my_mod
  my_samps[[i]] <- my_result$my_samp
}
all_results <- tibble(my_samps, my_mods, slopes)
```
---
class: inverse, center, middle

# End of Session 8.1