---
title: "COMPSCIX 415.2 Homework 9/Final"
author: "Shan Valleru"
date: "3/31/2019"
output: 
  html_document:
    self_contained: true
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

My Github repository for my assignments can be found at this URL: [https://github.com/svalleru/compscix-415-2-assignments](https://github.com/svalleru/compscix-415-2-assignments)

```{r chunk_name, warning=FALSE, message=FALSE}
library(tidyverse)
library(broom)
```

# Exercise 1 - Sampling Distributions, Functions and For Loops
```{r}
# sample size
samp_size <- 100
# set the rate parameter
samp_rate <- 1/10000

# take sample
rexp(n = samp_size, rate = samp_rate)
```

**STEP 1**

Write an R function that does the following:

- Takes a sample of size `samp_size` from this exponential distribution (`samp_size` is an input parameter for the function)
- Calculates the mean of that sample
- Calculates the standard deviation of that sample
- Returns the calculated mean and standard deviation as a list
```{r}
samp_fun <- function(samp_size, samp_rate) {

  sample <- rexp(n = samp_size, rate = samp_rate)
  samp_avg <- mean(sample)
  samp_std_dev <- sd(sample)
  
  stats <- list(samp_avg = samp_avg, samp_std_dev = samp_std_dev)
  return(stats)
}

```
**STEP 2**

Then write a loop that does this:

Runs the above function 1000 times, with `samp_size = 50` and `samp_rate = 1/10000`
Saves all of the sample means in a vector called `sample_means`, and all of the sample standard deviations in a vector called `sample_sds`
```{r}
iterations <- 1000
samp_size <- 50
samp_rate <- 1/1000

sample_means <- sample_sds <- c()

for (i in 1:iterations) {
  stats <- samp_fun(samp_size, samp_rate)
  sample_means[i] <- stats$samp_avg
  sample_sds[i] <- stats$samp_std_dev
}
```
**STEP 3**

Then

plot your sample means as a histogram
output the standard deviation of your sample means
calculate the theoretical standard error (σ=10000, n = sample size)
calculate the mean of the sample standard deviations and use this to calculate the empirical standard error
```{r}
mean_sd_values <- data.frame(mean = sample_means, sd = sample_sds)

# plot your sample means as a histogram
ggplot(data = mean_sd_values, mapping = aes(x = mean)) +
  geom_histogram(binwidth = 10)

# output the standard deviation of your sample means
sd(mean_sd_values$mean)

# calculate the theoretical standard error (σ=10000, n = sample size)
10000/sqrt(samp_size)

# calculate the mean of the sample standard deviations and use this to calculate the empirical standard error
mean_sample_sds <- mean(mean_sd_values$sd)
mean_sample_sds/sqrt(samp_size)
```

**STEP 4**

Repeat **STEP 2** and **STEP 3** using a sample size of `5000`.
```{r}
samp_size <- 5000
samp_rate <- 1/1000
sample_means <- sample_sds <- c()

for (i in 1:iterations) {
  stats <- samp_fun(samp_size, samp_rate)
  sample_means[i] <- stats$samp_avg
  sample_sds[i] <- stats$samp_std_dev
}

mean_sd_values <- data.frame(mean = sample_means, sd = sample_sds)
ggplot(data = mean_sd_values, mapping = aes(x = mean)) +
  geom_histogram(binwidth = 5)

# output the standard deviation of your sample means
sd(mean_sd_values$mean)

# calculate the theoretical standard error (σ=10000, n = sample size)
10000/sqrt(samp_size)

# calculate the mean of the sample standard deviations and use this to calculate the empirical standard error
mean_sample_sds <- mean(mean_sd_values$sd)
mean_sample_sds/sqrt(samp_size)

```

# Exercise 2 - Linear Regression 
Load the `train.csv` dataset into R and fit a regression model with:

- `y = SalePrice`
- Features: `LotArea`, `OverallQual`, and `ExterQual`

Answer these questions:

- Use the broom package to output the coefficients and the R-squared
- Interpret the coefficient on `LotArea`
- Interpret the coefficient on `ExterQualGd`
- Compare this model to the model we fit in HW 7 with `GrLivArea`, `OverallQual`, `Neighborhood`. Which is the better fitting model?
```{r}
training_data <- read_csv('/Users/svalleru/Desktop/UCB/COMPSCIX415.2-003 - Intro to DS/compscix-415-2-assignments/session9/train.csv')
model <- lm(formula = SalePrice ~ LotArea + OverallQual + ExterQual, data = training_data)
```

- Use the broom package to output the coefficients and the R-squared
```{r}
tidy(model)
glance(model)
```

- Interpret the coefficient on `LotArea`

When all other features are constant, 1 unit increase in `LotArea` increases `SalePrice` by \$1.45

- Interpret the coefficient on `ExterQualGd`

When all other features are constant, `SalePrice` for houses with `ExterQualGd`(Good exterior material quality) are on average \$71,529.49 lower than the houses with `ExterQualEx`(Excellent exteriot material quality)

- Compare this model to the model we fit in HW 7 with `GrLivArea`, `OverallQual`, `Neighborhood`. Which is the better fitting model?

Adjusted R-squared for this model is `0.694` which is lower than that of the model from HW7(`0.787`), so, HW7 model is a better fitting model

# Exercise 3 - AB Testing
Download the `ab_test_data.csv` file from Canvas. This file contains two columns: `version` and `conversion`. Each row is a visitor to a webpage. The `version` column tells us which version of the webpage the visitor saw, and the `conversion` column is a binary value and equals `1` if the visitor converted (`0` otherwise).

We want to perform an AB test on this data to see if the conversion rates are different for the two versions of the webpage.

Answer these questions:

- What proportion of visitors converted for each version of the webpage?
- Perform the AB test in R. What is the p-value for the AB test (hypothesis test of proportions)?

What proportion of visitors converted for each version of the webpage?
```{r}
ab_test_data <- read_csv('/Users/svalleru/Desktop/UCB/COMPSCIX415.2-003 - Intro to DS/compscix-415-2-assignments/session9/ab_test_data.csv')

conversion_stats <- 
  ab_test_data %>% 
  group_by(version) %>% 
  summarise(
    tatal_visited = n(), 
    total_converged = sum(conversion), 
    propotion = total_converged/tatal_visited)

conversion_stats
```
Version A: 4.15%

Version B: 10%

Perform the AB test in R. What is the p-value for the AB test (hypothesis test of proportions)?
```{r}
prop_test <- prop.test(conversion_stats$total_converged, conversion_stats$tatal_visited)
prop_test$p.value
```
p-value is way less than alpha(0.05), so, we can reject the default null hypothesis i.e., both versions have same conversion rate. In other words, conversion rates for both versions are significantly different from each other.


