---
title: "Introduction to Data Science"
subtitle: "Session 6.3"
author: "Robert Clements"
date: ""
output:
  xaringan::moon_reader:
    css: rc_css.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.align = "center", fig.asp=.5, message = FALSE, warning = FALSE)
```

```{r, message = FALSE, warning=FALSE, echo = FALSE}
library(tidyverse)
```

### Session 6.3 Outline

- Fundamental statistics concepts
  + Simpson's paradox
- Recoding variables
  + Factors
  + Dates and times
  + *R for Data Science Ch 15 & 16*

---
class: inverse, center, middle
# Simpson's paradox
---
### Simpson's paradox

... is when groups of data show one particular trend, but this trend is reversed when the groups are combined together.

Famous example: Berkeley admissions

From a study done in the 70's, we have:

1. admissions data by department (A-F)

2. gender of applicant

3. admissions decision
---
### Simpson's paradox

... is when groups of data show one particular trend, but this trend is reversed when the groups are combined together.


Famous example: Berkeley admissions

```{r}
ucb_ad <- read_csv(file = 'ucb_admiss.csv')
ucb_ad 
```
 
---
### Simpson's paradox

Here we plot the distribution of admission decisions overall **by gender**:

```{r}
ucb_ad %>% 
  ggplot() + geom_bar(aes(x = Gender, y = Freq, fill = Admit), stat = 'identity', position = 'fill') +
 labs(x = 'Gender', y = '', fill = 'Admission',
       title = 'Admit by gender') +
  theme_bw()
```
---
### Simpson's paradox

Here we plot the distribution of admission decisions **by gender and department**:

```{r}
ucb_ad %>% 
  ggplot() + geom_bar(aes(x = Gender, y = Freq, fill = Admit), stat = 'identity', position = 'fill') +
  facet_grid(~ Dept) +
  labs(x = 'Gender', y = '', fill = 'Admission',
       title = 'Admit by gender by department') +
  theme_bw()
```
---
### What happened?

Females tend to apply for more competitive departments.

Here we plot the distribution of applicants **by gender and department**:

```{r}
ucb_ad %>% 
  ggplot() + geom_bar(aes(x = Gender, y = Freq), stat = 'identity') +
  facet_grid(~ Dept) +
  labs(x = 'Gender', y = '', fill = 'Admission',
       title = 'Applicants by gender by department') +
  theme_bw()
```
---
### Recall this counterintuitive relationship

What can we do to try to reveal a more intuitive relationship using a third variable? We'll save that for later...

```{r}
diamonds %>% ggplot(aes(x = cut, y = price)) +
  geom_boxplot()
```

```{r, include = FALSE, eval = FALSE}
diamonds_2 <- diamonds %>% mutate(carat_fct = case_when(carat < 1 ~ '<1',
                                                        # carat >= .5 & carat < 1 ~ '<1',
                                                        carat >= 1 & carat < 1.5 ~ '<1.5',
                                                        carat >= 1.5 & carat < 2 ~ '<2',
                                                        carat >=2 ~ '>2'))
diamonds_2 %>% ggplot(aes(x = cut, y = price)) +
  geom_boxplot() +
  facet_grid(~carat_fct)
```
---
class: inverse, center, middle
# Recoding variables
---
### Recoding variables

Motivation:

- We know how to import/export our data (`read_***()` and `write_***()`)

- We know how to tidy our data (`spread()` and `gather()`)

- We know how to do aggregations (`summarize()`)

- We know how to subset and create new variables (`select()`, `filter()`, and `mutate()`)

- We know how to rearrange the data (`arrange()`)

---
### Recoding variables

But what do we do if we want to either: 

- convert from `numeric` to unordered `categorical` (`factor`),

- convert from `character` to unordered `factor`,

- convert from unordered `factor` to `ordinal`,

- rearrange the levels of a `factor`,

- recode a `categorical` variable with different categories,

- or convert a `character` to a `date`.
---
### More about `factors`

But first, what are `factors`?

A `factor` is a categorical or ordinal variable with a defined set of categories, or `levels`. By default, R will order the levels of a factor in alphabetical order.

```{r}
breed <- c('collie', 'collie', 'spaniel', 
           'poodle', 'poodle')
breed_fct <- factor(breed)
breed_fct

levels(breed_fct)
```
---
### More about `factors`

R treats `factors` differently than just `character` variables by making use of `levels`. Your factor can have levels that don't always appear in your variable.

```{r}
breed_fct <- factor(breed, levels = c('collie', 'poodle', 
                                      'spaniel', 'terrier', 
                                      'mutt'))

breed_fct
levels(breed_fct)
```
---
### More about `factors`

And your variable may have values that haven't been defined as a level.

```{r}
breed_new <- c('doberman', 'collie', 'collie', 'spaniel', 
               'poodle', 'poodle')
breed_new_fct <- factor(breed_new, levels = c('collie', 'poodle', 
                                              'spaniel', 'terrier', 
                                              'mutt'))

breed_new_fct
levels(breed_new_fct)
```
---
### Convert from `numeric` to `categorical`

Use `case_when()` from the `dplyr` package!!

`case_when()` takes a series of two-sided formulas, where:

- left hand side is a logical expression

- right hand side gives the value

Example:

```{r}
x <- 1:10
x
x_cat <- case_when(x < 3 ~ '<3',
                   x >= 3 & x < 6 ~ '3<=x<6',
                   x >= 6 ~ '>6')
```
---
### Convert from `numeric` to `categorical`

Use `case_when()` from the `dplyr` package!!

`case_when()` takes a series of two-sided formulas, where:

- left hand side is a logical expression

- right hand side gives the value

Example:

```{r}
x_cat
class(x_cat)
```
---
### Convert from `character` to unordered `factor`

`character` variables are **not** factors yet.

```{r}
class(x_cat)

x_fct <- factor(x_cat)
class(x_fct)
x_fct
```
---
### Convert from unordered `factor` to ordered `factor`

If you want ordinal variables, you have to be explicit

```{r}
class(x_fct)

x_fct <- factor(x_fct, ordered = TRUE)
class(x_fct)
x_fct
```
---
### Rearrange the levels of a `factor`

Use the `fct_relevel()` function from the `forcats` package (from the `tidyverse`)!!

`forcats` provides a whole bunch of convenience functions for dealing with `factors`.

```{r}
x_fct <- factor(x_fct, ordered = TRUE)
x_fct
levels(x_fct)
```
---
### Rearrange the levels of a `factor`

Use the `fct_relevel()` function from the `forcats` package (from the `tidyverse`)!!

```{r}
x_fct

x_fct <- fct_relevel(x_fct, '<3', '3<=x<6', '>6')

levels(x_fct)
```
---
### Pop Quiz?

Dealing with factors this way comes in handy during EDA and when building models with categorical variables. Why?


---
### Recode a `categorical` variable with different categories

Use `fct_recode()`

```{r}
x_fct_re <- fct_recode(x_fct,  'less than 3' = '<3',
                       'between 3 and 6' = '3<=x<6',
                       'other' = '>6')
x_fct_re
```
---
### Recode a `categorical` variable with different categories

We can also collapse categories with `fct_collapse()`

```{r}
x_fct_coll <- fct_collapse(x_fct, 'less than 3' = '<3',
                           'other' = c('3<=x<6', '>6'))
x_fct_coll
```

---
### Convert a `character` to a `date`

Dealing with dates can be a pain. Luckily, we can use the `lubridate` package.


Useful functions:

- `ymd()`, `ymd_hms()`, `dmy()`, `dmy_hms()`, `mdy()`: for parsing date-times

- `year()`, `month()`, `mday()`, `hour()`, `minute()`, `second()`: for extracting or setting components

- `with_tz()`, `force_tz()`: for dealing with time zones

---
### Convert a `character` to a `date`

Define a new character variable consisting of a date in a usual date format (ex. mm/dd/yyyy, yyyy-mm-dd, etc.)


```{r}
birthday <- c('6/16/1915') # John Tukey's birthday
birthday
class(birthday)
```
---
### Convert a `character` to a `date`

Parsing dates

```{r, message=FALSE, warning=FALSE}
library(lubridate)
birthday
birthday <- mdy(birthday)
birthday
class(birthday)
```
---
### Convert a `character` to a `date`

Parsing dates

```{r, message=FALSE, warning=FALSE}
birthday <- '1915-06-16 12:22:21.01'
birthday <- ymd_hms(birthday)
birthday
class(birthday)
```
---
### Convert a `character` to a `date`

Extracting and setting components

```{r, message=FALSE, warning=FALSE}
birthday

year(birthday)
hour(birthday) <- 5
birthday
```
---
### Convert a `character` to a `date`

Printing or changing time zones

Note: you can check `olson_time_zones()` for a list of time zones

```{r}
# Changes the time to the new time zone
with_tz(birthday, tzone = "America/Chicago")

# Keeps the time, but changes the time zone
force_tz(birthday, tzone = "America/Chicago")
```
---
### Pop Quiz

1. Enter your birthday into R, convert it to a Date object, and find the day of the week you were born using `weekdays()`.

2. How many days old are you? (Hint: use `Sys.Date()` to get today's date)
---
### If we have time, let's work on this problem...

What can we do to try to reveal a more intuitive relationship using a third variable? 

```{r}
diamonds %>% ggplot(aes(x = cut, y = price)) +
  geom_boxplot()
```

```{r, include = FALSE, eval = FALSE}
diamonds_2 <- diamonds %>% mutate(carat_fct = case_when(carat < 1 ~ '<1',
                                                        # carat >= .5 & carat < 1 ~ '<1',
                                                        carat >= 1 & carat < 1.5 ~ '<1.5',
                                                        carat >= 1.5 & carat < 2 ~ '<2',
                                                        carat >=2 ~ '>2'))
diamonds_2 %>% ggplot(aes(x = cut, y = price)) +
  geom_boxplot() +
  facet_grid(~carat_fct)
```
---
class: inverse, center, middle

# End of Session 6.3